Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-05 paper_jasa.tex
non-contaminated p-value plot removed
finished with Type III error section

\noindent with the possibility to use more than one. The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for thier choice. Reasons 1 or 3 would be indicative of Type III error. Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices. If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason.

It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2. Figure \ref{fig:choice_reason} examines this. For observers who chose the actual data plot, the proportions for the three reasons are shown. It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third. \green{What about other??} \blue{Not many people chose other as we see in Figure \ref{fig:choice_reason}.}

@@

\begin{figure}[hbtp]
   \centering
%       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
%       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
       \caption{Reasons for choosing the plot in a lineup for experiment 3. Reason 12 means both reasons 1 and 2 are chosen. Number of subjects choosing that reason are shown on the right hand side. Notice that 345 people choose only reason (3) "Clustering visible", out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error). Those who chose reason 123 got the actual plot about 97\% times.}
       \label{fig:choice_reason}
\end{figure}

>>>>

\noindent with the possibility to use more than one. The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice. Reasons 1 or 3 would be indicative of Type III error. Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices. If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot. This is not what we see. At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons. The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate. The size of the point corresponds to the number of subjects putting this as the reaosn. Subjects that chose all three reasons almost always chose the actual data plot. This was followed by using 1 and 3, and then 1 and 4. The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted. ``Visible trend'' comes in third. This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.

%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2. Figure \ref{fig:choice_reason} examines this. For observers who chose the actual data plot, the proportions for the three reasons are shown. It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third. \green{What about other??} \blue{Not many people chose other as we see in Figure \ref{fig:choice_reason}.}

@@

\begin{figure}[hbtp]
   \centering
%       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
%       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
       \caption{Reasons for choosing the plot in a lineup for experiment 3. Reason 12 means both reasons 1 and 2 are chosen. Number of subjects choosing that reason are shown on the right hand side. Notice that 345 people choose only reason (3) "Clustering visible", out of whom 64\% got the actual plot (proportion correct = 0.64), which indicates there are number of people who choose the right plot for wrong reason (Type-III error). Those who chose reason 123 got the actual plot about 97\% times.}
       \label{fig:choice_reason}
\end{figure}

For visual inference, making a Type III, is not actually a problem. It is only a possibility in this ecperiment because we are working with known structure. In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis. This does show the importance of the qualitative answers that observers give for their choices.
